A numerical study of variational data assimilation for the wave equation in heterogeneous media

### 2025-06-23 tex/main.tex
Reproduction (WIP)

\author[1]{\fnm{Erik} \sur{Burman}}\email{e.burman@ucl.ac.uk}
\author[1]{\fnm{Janosch} \sur{Preuss}}\email{j.preuss@ucl.ac.uk}
\author[2]{\fnm{Tim} \sur{van Beeck}}\email{t.beeck@math.uni-goettingen.de}

@@

\affil[2]{\orgdiv{Institute for Numerical and Applied Mathematics}, \orgname{University of G\"{o}ttingen}, \orgaddress{\street{Lotzestr. 16-18}, \city{G\"{o}ttingen}, \postcode{37083}, \country{Germany}}}

@@

\section{Introduction}
\noindent In this article, we consider the discretization of a data assimilation problem for the wave equation in heterogeneous media. In the following, let $\Omega \subset \mathbb{R}^d$ be a bounded domain with smooth boundary $\partial \Omega$. We assume that $\Omega$ is split by a smooth $(d-1)$-dimensional 
interface $\Gamma$ into two disjoint components, i.e.\ $\Omega \setminus \Gamma = \Omega_1 \cup 
\Omega_2$ and $\Omega_1 \cap \Omega_2 = \emptyset$.
 We consider a scalar wave speed
\[
 c(x) := \begin{cases}
       c_1(x),  & x \in \Omega_1, \\
        c_2(x), & x \in \Omega_2,
    \end{cases}
    \quad 
    \text{ with } c_i \in C^\infty(\Omega_i), \; i = 1,2,
\]
 %$c(x) = \mathbbm{1}_{\Omega_1} c_1(x) + \mathbbm{1}_{\Omega_2} c_2(x)$, $c_i \in C^\infty(\Omega_i)$, $i = 1,2$,
which may be discontinuous across the interface $\Gamma$, but constant in time. For a final time $T > 0$, we consider the space-time domain $Q = (0,T) \times \Omega$ with lateral boundary $\Sigma \coloneqq (0,T) \times \partial \Omega$. For any nonempty subset $\omega \subset \Omega$, we set $\omega_T \coloneqq (0,T) \times \omega$. In the following, let $\wop$ be the wave operator with discontinuous wave-speed defined as 
\begin{equation*}
    \wop u := \partial_t^2 - \div(c^2 \nabla u).
\end{equation*}
In particular, we write $\square_1$ if $c_i = 1$, $i = 1,2$, i.e. if we are considering the case of homogeneous media. 
Then, we consider the problem: find $u : Q \rightarrow \mathbb{R}$ such that 
\begin{equation}\label{eq:waveEquation}
        (\wop u, u) = (0,0) \quad \text{ in } Q \times \Sigma,
\end{equation}
along with the natural interface conditions 
\begin{equation}\label{eq:IF-conditions}
	\jump{u}_{\Gamma} = 0 \text{ on } \Sigma, \qquad 
	\quad \jump{c^2 \nabla u}_{\Gamma} \cdot \mathbf{n}_{\Gamma} = 0 \text{ on } \Sigma, 
\end{equation}
where $\mathbf{n}_{\Gamma}$ is the normal vector on $\Gamma$ pointing into $\Omega_+$ 
and the jump across $\Gamma$ is defined\textsuperscript{1}\footnotemark \footnotetext{\textsuperscript{1}And anologously for vector-valued functions.} as $\jump{u}_{\Gamma} = u_{+}-u_{-}$ with $u_{\pm}$ being the traces 
of $u|_{\Omega_{\pm}}$ on $\Gamma$. \par 
\noindent To ensure that the problem is well posed, one usually prescribes initial data
\begin{equation}\label{eq:initialData}
    (u,\dt u) \vert_{t = 0} = (u_0,u_1) \text{ in } \Omega, \tag{IVP} 
\end{equation}
where $u_0 \in H^1(Q)$ and $u_1 \in L^2(Q)$ are given. For a detailed analysis of the well-posedness of the problem \eqref{eq:waveEquation}-\eqref{eq:IF-conditions}+\eqref{eq:initialData} in heterogeneous media we refer to \cite{StolkPhD}. 
Here we consider a data assimilation problem instead, i.e. we assume that the initial data is unknown and that there are given measurements $u_{\omega} \in L^2(\omega_T)$ which are prescribed in a data domain $\omega_T$, $\omega \subset \Omega$, $\omega \not = \emptyset$:
\begin{equation}\label{eq:dataMatch}
    u = u_{\omega} \text{ in } \omega_T. \tag{DA}
\end{equation}

@@

\subsection{Discretization of the space-time domain}\label{sec:spaceTimeDiscretization}
Let $\hat{\mathcal{T}}_h$ be a quasi-uniform triangulation with mesh size $h$ such that $\hat{\Omega}:= \cup_{K \in \hat{\mathcal{T}}_h} K $ contains $\Omega$, i.e.\ $\Omega \subset \hat{\Omega}$. Setting $\mathcal{T}_h := \{ K \cap \Omega \mid K \in  \hat{\mathcal{T}}_h \}$ then ensures that $\Omega = \cup_{K \in \mathcal{T}_h} K$ holds true.  
We recall from  \cite[Sec. 4.2]{BFMO21control} that the triangulation $\{ \hat{\mathcal{T}}_h \mid h > 0 \}$ can be constructed such that for $h$ small enough the following continuous trace inequality holds:  For all $v \in [H^1(K)]^d$ and $K \in \mathcal{T}_h$, we have
\begin{equation}\label{eq:traceInequality}
    \Vert v \Vert_{[L^2(\partial K)]^d} \le C \left(h^{-1/2} \Vert v \Vert_{[L^2(K)]^d} + h^{1/2} \Vert \nabla v \Vert_{[L^2(K)]^d} \right).  
\end{equation}
We assume in what follows additionally that the triangulation $\mathcal{T}_h$ fits the interface $\Gamma$ and the data domain $\omega$. An extension to allow for unfitted interfaces could be possible utilizing techniques we introduced for a stationary problem in \cite{BP23arxiv}, but is certainly beyond the scope of this article.

@@

\begin{figure}[!htbp]
    \begin{center}
        \begin{tikzpicture}[scale=0.8]
            \begin{groupplot}[%
                group style={%
                group size=2 by 2,
                horizontal sep=1.25cm,
                vertical sep=2cm,
                },
            ymajorgrids=true,
            grid style=dashed,
            %ymin = 1e-3, ymax = 0.5e1,
            ]    
            \nextgroupplot[width=8cm,height=6.2cm,domain=1:4,xmode=linear,ymode=log, ylabel={}, xlabel={$L$}, title={$T=0.5$, $\Vert (u-L_{\Delta t} \ul_1) \Vert_{L^\infty(0,T;L^2(\Omega_R))}$}, %cycle list name=, 
            legend style={legend columns=4, draw=none,nodes={scale=.8}},legend to name=named,
            ]

@@

@@

            \addplot[gray, dashed, domain=1:4] {4.5*(1/2^(2))^(x-0.9)};
            \addplot[gray, dotted, domain=1:4] {1.0*(1/2^(3))^(x-0.9)};
            %\node [draw=none] at (axis description cs:0.30,0.87) {\color{gray}\footnotesize $\!\!\mathcal{O}(h^{2})$};
            %\node [draw=none] at (axis description cs:0.65,0.2) {\color{gray}\footnotesize $\!\!\mathcal{O}(h^{3})$};
            %\legend{$k=2 $,, $k = 3$,}
            \nextgroupplot[width=8cm,height=6.2cm,domain=1:4,xmode=linear,ymode=log, ylabel={}, xlabel={$L$}, title={$T = 0.1$, $\Vert (u-L_{\Delta t} \ul_1) \Vert_{L^\infty(0,T;L^2(\Omega))}$}, %cycle list name=, 
            legend pos=south west, %yticklabels={,,},
            ]

@@

            %\addplot[gray, dashed, domain=1:4] {15*(1/2^(2))^(x-0.9)};
            \addplot[gray, dashed, domain=1:4] {0.09*(1/2^(2))^(x-0.9)};
            %\addplot[gray, dashed, domain=1:4] {7.5*(1/2^(3))^(x-0.9)};
            \addplot[gray, dotted, domain=1:4] {0.05*(1/2^(3))^(x-0.9)};
            \addplot[gray,dashed] table {
                1 0.009999968541169805
                2 0.006666645694113204
                3 0.004999984270584902
                4 0.003999987416467922

            };
            %\node [draw=none] at (axis description cs:0.75,0.45) {\color{gray}\footnotesize $\!\!\mathcal{O}(h^{2})$};
            %\node [draw=none] at (axis description cs:0.45,0.35) {\color{gray}\footnotesize $\!\!\mathcal{O}(h^{3})$};
            \node [draw=none] at (axis description cs:0.7,0.75) {\color{gray}\footnotesize $\!\!\mathcal{O}(\vert \log(h) \vert^{-1})$};

@@

            \nextgroupplot[width=8cm,height=6.2cm,domain=1:4,xmode=linear,ymode=log, ylabel={}, xlabel={$L$}, title={$T = 0.1$, $\Vert \partial_t (u-L_{\Delta t} \ul_1) \Vert_{L^2(0,T;L^2(\Omega))}$}, %cycle list name=, 
            legend pos=south west, %yticklabels={,,},
            ]
            %L2-L2-u_t (k = 2)
            \addplot+[discard if not={order}{2}, discard if not={contrast}{2.5},line width=1.1pt,color=orange,mark=square*] table [x=L, y=L2-L2-error-u_t, col sep=comma] {../dataNew/simplexExact_singleContrast_T0.1.csv};
            \addplot+[discard if not={order}{2}, discard if not={contrast}{2.5},line width=1.1pt,dashed,color=orange,mark=square*] table [x=L, y=bestapprox-L2-L2-error-u_t, col sep=comma] {../dataNew/simplexExact_singleContrast_T0.1.csv};

@@

            \addplot[gray, dashed, domain=1:4] {0.5*(1/2^(2))^(x-0.9)};
            %\addplot[gray, dashed, domain=1:4] {1.0*(1/2^(2))^(x-0.9)};
            \addplot[gray, dotted, domain=1:4] {0.2*(1/2^(3))^(x-0.9)};
            %\addplot[gray, dashed, domain=1:4] {0.5*(1/2^(3))^(x-0.9)};
            \addplot[gray,dashed] table {
                1 0.09999968541169805
                2 0.06666645694113204
                3 0.04999984270584902
                4 0.03999987416467922

            };
            %\node [draw=none] at (axis description cs:0.80,0.40) {\color{gray}\footnotesize $\!\!\mathcal{O}(h^{2})$};
            %\node [draw=none] at (axis description cs:0.65,0.2) {\color{gray}\footnotesize $\!\!\mathcal{O}(h^{3})$};
            \node [draw=none] at (axis description cs:0.7,0.75) {\color{gray}\footnotesize $\!\!\mathcal{O}(\vert \log(h) \vert^{-1})$};
            %\legend{$k=2 $,, $k = 3$,}
            \end{groupplot}
        \end{tikzpicture}
        \pgfplotslegendfromname{named}
    \end{center}
    \caption{For the approximation of \eqref{eq:1D:exact:simple} with polynomial degree $k \in \{2,3\}$ and contrast $c_1 = 2.5$ we measure the errors $\Vert u-L_{\Delta t} \ul_1 \Vert_{L^\infty(0,T;L^2(\Omega_R))}$ (left) and $\Vert \partial_t (u-L_{\Delta t} \ul_1) \Vert_{L^2(0,T;L^2(\Omega_R))}$ (right) and compare to the respective error of the $L^2$-bestapproximation (dashed). We consider the final time $T = 0.5$ (top) and $T = 0.1$ (bottom). The gray dashed lines indicate the observed convergence rates. \textbf{adjust log lines..., shared legend...}}
    \label{fig:jumpCoefs:contrast:2.5}
  \end{figure}

  \begin{figure}[!htbp]
    \begin{center}
        \begin{tikzpicture}[scale=0.85]
            \begin{groupplot}[%
                group style={%
                group size=2 by 1,
                horizontal sep=1.0cm,
                vertical sep=2cm,
                },
            ymajorgrids=true,
            grid style=dashed,
            %ymin = 1e-3, ymax = 0.5e1,
            ]    
            \nextgroupplot[width=8cm,height=6.2cm,domain=1:5,xmode=linear,ymode=log, ylabel={}, xlabel={$c_1$}, title={$\Vert \partial_t (u-L_{\Delta t} \ul_1) \Vert_{L^2(0,T;L^2(\Omega_R))}$}, %cycle list name=, 
            legend style={legend columns=4, draw=none,nodes={scale=.8}},legend to name=named, %yticklabels={,,},
            %ymin=1e-7,ymax=1e1,
            ]
            %\addplot+[line width=1.1pt,color=teal,mark=*] table [x=contrast, y=L2-L2-error-u_t, col sep=comma]{../data/simpleExact_1D_jumpingCoefs_k3.csv};
            \addplot+[line width=1.1pt,color=teal,mark=*] table [x=contrast, y=L2-L2-error-u_t, col sep=comma]{../dataNew/simplexExact_MultContrast_T0.5.csv};
            \addplot+[line width=1.1pt,color=teal,mark=*,dashed] table [x=contrast, y=bestapprox-L2-L2-error-u_t, col sep=comma]{../dataNew/simplexExact_MultContrast_T0.5.csv};
            \addplot+[line width=1.1pt,color=orange,mark=diamond*] table [x=contrast, y=L2-L2-error-u_t, col sep=comma]{../dataNew/simplexExact_MultContrast_TAdap.csv};
            \addplot+[line width=1.1pt,color=orange,mark=diamond*,dashed] table [x=contrast, y=bestapprox-L2-L2-error-u_t, col sep=comma]{../dataNew/simplexExact_MultContrast_TAdap.csv};

@@

            \legend{$T = 0.5$, $L^2$-best approx. ($T = 0.5$), $T$ variable, $L^2$-best approx. ($T$ variable)}
            \nextgroupplot[width=8cm,height=6.2cm,domain=1:5,xmode=linear,ymode=log, ylabel={}, xlabel={$c_1$}, title={$\Vert u-L_{\Delta t} \ul_1 \Vert_{L^\infty(0,T;L^2(\Omega_R))}$}, %cycle list name=, 
            legend pos=south east, %yticklabels={,,},
            %ymin=1e-7,ymax=1e1,
            ]
            \addplot+[line width=1.1pt,color=teal,mark=*] table [x=contrast, y=L-infty-L2-error-u, col sep=comma]{../dataNew/simplexExact_MultContrast_T0.5.csv};
            \addplot+[line width=1.1pt,color=teal,mark=*,dashed] table [x=contrast, y=bestapprox-L-infty-L2-error-u, col sep=comma]{../dataNew/simplexExact_MultContrast_T0.5.csv};

@@

                \addplot+[discard if not={L}{4},discard if not={contrast}{2.5},line width=3pt,mark=None,color=black] table [x=x, y=y, col sep=comma] {../dataNew/plots/simpleExact_ApproxPlot_T0.5.csv};

@@

                \legend{exact,$L=1$, $L=2$, $L=3$, $L=4$}
                \nextgroupplot[width=9cm,height=7cm,domain=0:1,xmode=linear,ymode=linear, xlabel={$x$}, ylabel={}, title={$u$ vs. $u_h$ ($c_1 = 5.5$)}, %cycle list name=paulcolors, 
                legend pos=north west, %yticklabels={,,},
                ymin = -2.25, ymax = 2.25,
                ]
                %\draw[fill=gray!40,draw=none,opacity=0.5] (0.25,-3) -- (0.25,3) -- (0,3) -- (0,-3) -- cycle;
                %\draw[fill=gray!40,draw=none,opacity=0.5] (0.75,-3) -- (0.75,3) -- (1.0,3) -- (1.0,-3) -- cycle;
                %\addplot+[discard if not={contrast}{1.0},line width=2.5pt,mark=None,color=black] table [x=x, y=y, col sep=comma] {../../wave_data_assimilation/data/jumpCoefs/exact_plot_data.csv};
                %\addplot+[discard if not={contrast}{5.5},line width=3.0pt,mark=None,color=black] table [x=x, y=y, col sep=comma] {../data/exact_plot_data_1D_simple.csv};

                \addplot+[discard if not={L}{4},discard if not={contrast}{5.5},line width=3pt,mark=None,color=black] table [x=x, y=y, col sep=comma] {../dataNew/plots/simpleExact_ApproxPlot_T0.5.csv};

@@

            \end{groupplot}
        \end{tikzpicture}
        \pgfplotslegendfromname{named}
    \end{center}
    \caption{Exact solution \eqref{eq:1D:exact:simple} and approximations with $k = 3$ on refinement levels $L \in \{1,2,3,4 \}$ for contrasts $c_1 = 2.5$ (left) and $c_1= 5.5$ (right). The scale in the right plot indicates the mesh size $h = 1/2^{L+1}$ for the different refinement levels.v \textbf{generate the data for the exact plots, at which time are we?}}
    \label{fig:jumpCoefs:quality}
  \end{figure}

@@

                \addplot+[discard if not={L}{1},discard if not={contrast}{7.5},line width=3.0pt,mark=None,color=yellow] table [x=x, y=y, col sep=comma] {../dataNew/plots/MultipleJumpsExact_ApproxPlot_T0.5.csv};

                %\addplot+[line width=1.0pt,mark=None,color=orange] table [x=x, y=y, col sep=comma] {../data/multipleJumps_approx_plot_data_WPFalse_contrast7.5_k3_uniform.csv};
                \addplot+[line width=1.0pt,mark=None,color=blue!75!white] table [x=x, y=y, col sep=comma] {../data/multipleJumps_approx_plot_data_WPFalse_contrast7.5_k3_uniform_T0.5_L4.csv};
                \addplot+[line width=1.0pt,mark=None,color=orange] table [x=x, y=y, col sep=comma] {../data/multipleJumps_approx_plot_data_WPFalse_contrast7.5_k3_uniform_T0.5_L3.csv};

>>>>

\author[1]{\fnm{Erik} \sur{Burman}}\email{e.burman@ucl.ac.uk}
\author[2]{\fnm{Janosch} \sur{Preuss}}\email{janosch.preuss@inria.fr}
\author[3]{\fnm{Tim} \sur{van Beeck}}\email{t.beeck@math.uni-goettingen.de}

@@

\affil[2]{\orgdiv{MAKUTU}, \orgname{Inria}, \orgaddress{\street{...}, \city{...}, \postcode{...}, \country{France}}}

\affil[3]{\orgdiv{Institute for Numerical and Applied Mathematics}, \orgname{University of G\"{o}ttingen}, \orgaddress{\street{Lotzestr. 16-18}, \city{G\"{o}ttingen}, \postcode{37083}, \country{Germany}}}

@@

\section{Introduction}
\todo[inline]{@JP: please update affiliation}
\noindent We consider the discretization of a data assimilation problem for the wave equation in heterogeneous media. In the following, let $\Omega \subset \mathbb{R}^d$ be a bounded domain with smooth boundary $\partial \Omega$. We assume that $\Omega$ is split by a smooth $(d-1)$-dimensional 
interface $\Gamma$ into two disjoint components, i.e.\ $\Omega \setminus \Gamma = \Omega_1 \cup 
\Omega_2$ and $\Omega_1 \cap \Omega_2 = \emptyset$.
 We consider a scalar wave speed
\[
 c(x) := \begin{cases}
       c_1(x),  & x \in \Omega_1, \\
        c_2(x), & x \in \Omega_2,
    \end{cases}
    \quad 
    \text{ with } c_i \in C^\infty(\Omega_i), \; i = 1,2,
\]
 %$c(x) = \mathbbm{1}_{\Omega_1} c_1(x) + \mathbbm{1}_{\Omega_2} c_2(x)$, $c_i \in C^\infty(\Omega_i)$, $i = 1,2$,
which may be discontinuous across the interface $\Gamma$, but constant in time. For a final time $T > 0$, we consider the space-time domain $Q = (0,T) \times \Omega$ with lateral boundary $\Sigma \coloneqq (0,T) \times \partial \Omega$. For any nonempty subset $\omega \subset \Omega$, we set $\omega_T \coloneqq (0,T) \times \omega$. In the following, let $\wop$ be the wave operator with discontinuous wave-speed defined as 
\begin{equation*}
    \wop u := \partial_t^2 - \div(c^2 \nabla u).
\end{equation*}
In particular, we write $\square_1$ if $c_i = 1$, $i = 1,2$, i.e. if we are considering the case of homogeneous media. 
Then, we consider the problem: find $u : Q \rightarrow \mathbb{R}$ such that 
\begin{equation}\label{eq:waveEquation}
        (\wop u, u) = (0,0) \quad \text{ in } Q \times \Sigma,
\end{equation}
along with the natural interface conditions 
\begin{equation}\label{eq:IF-conditions}
	\jump{u}_{\Gamma} = 0 \text{ on } \Sigma, \qquad 
	\quad \jump{c^2 \nabla u}_{\Gamma} \cdot \mathbf{n}_{\Gamma} = 0 \text{ on } \Sigma, 
\end{equation}
where $\mathbf{n}_{\Gamma}$ is the normal vector on $\Gamma$ pointing into $\Omega_+$ 
and the jump across $\Gamma$ is defined\textsuperscript{1}\footnotemark \footnotetext{\textsuperscript{1}And anologously for vector-valued functions.} as $\jump{u}_{\Gamma} = u_{+}-u_{-}$ with $u_{\pm}$ being the traces 
of $u|_{\Omega_{\pm}}$ on $\Gamma$. \par 
\noindent To ensure that the problem is well posed, one usually prescribes initial data
\begin{equation}\label{eq:initialData}
    (u,\dt u) \vert_{t = 0} = (u_0,u_1) \text{ in } \Omega, \tag{IVP} 
\end{equation}
where $u_0 \in H^1(Q)$ and $u_1 \in L^2(Q)$ are given. For a detailed analysis of the well-posedness of the problem \eqref{eq:waveEquation}-\eqref{eq:IF-conditions}+\eqref{eq:initialData} in heterogeneous media we refer to \cite{StolkPhD}. 
Here we consider a data assimilation problem instead, i.e. we assume that the initial data is unknown and that there are given measurements $u_{\omega} \in L^2(\omega_T)$ which are prescribed in a data domain $\omega_T$, $\omega \subset \Omega$, $\omega \not = \emptyset$:
\begin{equation}\label{eq:dataMatch}
    u = u_{\omega} \text{ in } \omega_T. \tag{DA}
\end{equation}

@@

\subsection{Discretization of the space-time domain}\label{sec:spaceTimeDiscretization}
Let $\hat{\mathcal{T}}_h$ be a quasi-uniform triangulation with mesh size $h$ such that $\hat{\Omega}:= \cup_{K \in \hat{\mathcal{T}}_h} K $ contains $\Omega$, i.e.\ $\Omega \subset \hat{\Omega}$. Setting $\mathcal{T}_h := \{ K \cap \Omega \mid K \in  \hat{\mathcal{T}}_h \}$ then ensures that $\Omega = \cup_{K \in \mathcal{T}_h} K$ holds true.  
We recall from  \cite[Sec. 4.2]{BFMO21control} that the triangulation $\{ \hat{\mathcal{T}}_h \mid h > 0 \}$ can be constructed such that for $h$ small enough the following continuous trace inequality holds:  For all $v \in [H^1(K)]^d$ and $K \in \mathcal{T}_h$, we have
\begin{equation}\label{eq:traceInequality}
    \Vert v \Vert_{[L^2(\partial K)]^d} \le C \left(h^{-1/2} \Vert v \Vert_{[L^2(K)]^d} + h^{1/2} \Vert \nabla v \Vert_{[L^2(K)]^d} \right).  
\end{equation}
We assume in what follows additionally that the triangulation $\mathcal{T}_h$ fits the interface $\Gamma$ and the data domain $\omega$. An extension to allow for unfitted interfaces could be possible utilizing techniques we introduced for a stationary problem in \cite{BP25}, but is certainly beyond the scope of this article.

@@

\begin{figure}[!htbp]
    \begin{center}
        \begin{tikzpicture}[scale=0.8]
            \begin{groupplot}[%
                group style={%
                group size=2 by 2,
                horizontal sep=1.25cm,
                vertical sep=2cm,
                },
            ymajorgrids=true,
            grid style=dashed,
            %ymin = 1e-3, ymax = 0.5e1,
            ]    
            \nextgroupplot[width=8cm,height=6.2cm,domain=1:4,xmode=linear,ymode=log, ylabel={}, xlabel={$L$}, title={$T=0.5$, $\Vert (u-L_{\Delta t} \ul_1) \Vert_{L^\infty(0,T;L^2(\Omega_R))}$}, %cycle list name=,
            xtick={1,2,3,4}, 
            legend style={legend columns=4, draw=none,nodes={scale=.8}},legend to name=named,
            ]

@@

            \nextgroupplot[width=8cm,height=6.2cm,domain=1:4,xmode=linear,ymode=log, ylabel={}, xlabel={$L$}, title={$T = 0.5$, $\Vert \partial_t (u-L_{\Delta t} \ul_1) \Vert_{L^2(0,T;L^2(\Omega_R))}$}, %cycle list name=, 
            xtick={1,2,3,4},
            legend pos=south west, %yticklabels={,,},
            ]

@@

            \addplot[gray, dashed, domain=1:4] {4.5*(1/2^(2))^(x-0.9)};
            \addplot[gray, dotted, domain=1:4] {1.0*(1/2^(3))^(x-0.9)};
            %\node [draw=none] at (axis description cs:0.30,0.87) {\color{gray}\footnotesize $\!\!\mathcal{O}(h^{2})$};
            %\node [draw=none] at (axis description cs:0.65,0.2) {\color{gray}\footnotesize $\!\!\mathcal{O}(h^{3})$};
            %\legend{$k=2 $,, $k = 3$,}
            \nextgroupplot[width=8cm,height=6.2cm,domain=1:4,xmode=linear,ymode=log, ylabel={}, xlabel={$L$}, title={$T = 0.1$, $\Vert (u-L_{\Delta t} \ul_1) \Vert_{L^\infty(0,T;L^2(\Omega))}$}, %cycle list name=, 
            xtick={1,2,3,4},
            legend pos=south west, %yticklabels={,,},
            ]

@@

            %\addplot[gray, dashed, domain=1:4] {15*(1/2^(2))^(x-0.9)};
            \addplot[gray, dashed, domain=1:4] {0.09*(1/2^(2))^(x-0.9)};
            %\addplot[gray, dashed, domain=1:4] {7.5*(1/2^(3))^(x-0.9)};
            \addplot[gray, dotted, domain=1:4] {0.05*(1/2^(3))^(x-0.9)};
            %\addplot[gray,dashed] table {
            %    1 0.009999968541169805
            %    2 0.006666645694113204
            %    3 0.004999984270584902
            %    4 0.003999987416467922
            %};
            \addplot[gray,dashdotted] table {
                1 1.609999968541169805
                2 1.606666645694113204
                3 1.604999984270584902
                4 1.603999987416467922
            };
            %\node [draw=none] at (axis description cs:0.75,0.45) {\color{gray}\footnotesize $\!\!\mathcal{O}(h^{2})$};
            %\node [draw=none] at (axis description cs:0.45,0.35) {\color{gray}\footnotesize $\!\!\mathcal{O}(h^{3})$};
            \node [draw=none] at (axis description cs:0.7,0.94) {\color{gray}\footnotesize $\!\!\mathcal{O}(\vert \log(h) \vert^{-1})$};

@@

            \nextgroupplot[width=8cm,height=6.2cm,domain=1:4,xmode=linear,ymode=log, ylabel={}, xlabel={$L$}, title={$T = 0.1$, $\Vert \partial_t (u-L_{\Delta t} \ul_1) \Vert_{L^2(0,T;L^2(\Omega))}$}, %cycle list name=, 
            xtick={1,2,3,4},
            legend pos=south west, %yticklabels={,,},
            ]
            %L2-L2-u_t (k = 2)
            \addplot+[discard if not={order}{2}, discard if not={contrast}{2.5},line width=1.1pt,color=orange,mark=square*] table [x=L, y=L2-L2-error-u_t, col sep=comma] {../dataNew/simplexExact_singleContrast_T0.1.csv};
            \addplot+[discard if not={order}{2}, discard if not={contrast}{2.5},line width=1.1pt,dashed,color=orange,mark=square*] table [x=L, y=bestapprox-L2-L2-error-u_t, col sep=comma] {../dataNew/simplexExact_singleContrast_T0.1.csv};

@@

            \addplot[gray, dashed, domain=1:4] {0.5*(1/2^(2))^(x-0.9)};
            %\addplot[gray, dashed, domain=1:4] {1.0*(1/2^(2))^(x-0.9)};
            \addplot[gray, dotted, domain=1:4] {0.2*(1/2^(3))^(x-0.9)};
            %\addplot[gray, dashed, domain=1:4] {0.5*(1/2^(3))^(x-0.9)};
            %\addplot[gray,dashed] table {
            %    1 0.09999968541169805
            %    2 0.06666645694113204
            %    3 0.04999984270584902
            %    4 0.03999987416467922
            %};
            \addplot[gray,dashdotted] table {
                1 7.59999968541169805
                2 7.56666645694113204
                3 7.54999984270584902
                4 7.53999987416467922

            };
            %\node [draw=none] at (axis description cs:0.80,0.40) {\color{gray}\footnotesize $\!\!\mathcal{O}(h^{2})$};
            %\node [draw=none] at (axis description cs:0.65,0.2) {\color{gray}\footnotesize $\!\!\mathcal{O}(h^{3})$};
            \node [draw=none] at (axis description cs:0.7,0.94) {\color{gray}\footnotesize $\!\!\mathcal{O}(\vert \log(h) \vert^{-1})$};
            %\legend{$k=2 $,, $k = 3$,}
            \end{groupplot}
        \end{tikzpicture}
        \pgfplotslegendfromname{named}
    \end{center}
    \caption{For the approximation of \eqref{eq:1D:exact:simple} with polynomial degree $k \in \{2,3\}$ and contrast $c_1 = 2.5$ we measure the errors $\Vert u-L_{\Delta t} \ul_1 \Vert_{L^\infty(0,T;L^2(\Omega_R))}$ (left) and $\Vert \partial_t (u-L_{\Delta t} \ul_1) \Vert_{L^2(0,T;L^2(\Omega_R))}$ (right) and compare to the respective error of the $L^2$-best approximation (dashed). We consider the final time $T = 0.5$ (top) and $T = 0.1$ (bottom). The gray dashed lines indicate the observed convergence rates.}
    \label{fig:jumpCoefs:contrast:2.5}
  \end{figure}

  \begin{figure}[!htbp]
    \begin{center}
        \begin{tikzpicture}[scale=0.85]
            \begin{groupplot}[%
                group style={%
                group size=2 by 1,
                horizontal sep=1.0cm,
                vertical sep=2cm,
                },
            ymajorgrids=true,
            grid style=dashed,
            %ymin = 1e-3, ymax = 0.5e1,
            ]    
            \nextgroupplot[width=8cm,height=6.2cm,domain=1:5,xmode=linear,ymode=log, ylabel={}, xlabel={$c_1$}, title={$\Vert \partial_t (u-L_{\Delta t} \ul_1) \Vert_{L^2(0,T;L^2(\Omega_R))}$}, %cycle list name=, 
            legend style={legend columns=4, draw=none,nodes={scale=.8}},legend to name=named, %yticklabels={,,},
            xtick={1,1.5,2,2.5,3,3.5,4,4.5},
            %ymin=1e-7,ymax=1e1,
            ]
            %\addplot+[line width=1.1pt,color=teal,mark=*] table [x=contrast, y=L2-L2-error-u_t, col sep=comma]{../data/simpleExact_1D_jumpingCoefs_k3.csv};
            \addplot+[line width=1.1pt,color=teal,mark=*] table [x=contrast, y=L2-L2-error-u_t, col sep=comma]{../dataNew/simplexExact_MultContrast_T0.5.csv};
            \addplot+[line width=1.1pt,color=teal,mark=*,dashed] table [x=contrast, y=bestapprox-L2-L2-error-u_t, col sep=comma]{../dataNew/simplexExact_MultContrast_T0.5.csv};
            \addplot+[line width=1.1pt,color=orange,mark=diamond*] table [x=contrast, y=L2-L2-error-u_t, col sep=comma]{../dataNew/simplexExact_MultContrast_TAdap.csv};
            \addplot+[line width=1.1pt,color=orange,mark=diamond*,dashed] table [x=contrast, y=bestapprox-L2-L2-error-u_t, col sep=comma]{../dataNew/simplexExact_MultContrast_TAdap.csv};

@@

            \legend{$T = 0.5$, $L^2$-best approx. ($T = 0.5$), $T$ variable, $L^2$-best approx. ($T$ variable)}
            \nextgroupplot[width=8cm,height=6.2cm,domain=1:5,xmode=linear,ymode=log, ylabel={}, xlabel={$c_1$}, title={$\Vert u-L_{\Delta t} \ul_1 \Vert_{L^\infty(0,T;L^2(\Omega_R))}$}, %cycle list name=, 
            legend pos=south east, %yticklabels={,,},
            %ymin=1e-7,ymax=1e1,
            xtick={1,1.5,2,2.5,3,3.5,4,4.5},
            ]
            \addplot+[line width=1.1pt,color=teal,mark=*] table [x=contrast, y=L-infty-L2-error-u, col sep=comma]{../dataNew/simplexExact_MultContrast_T0.5.csv};
            \addplot+[line width=1.1pt,color=teal,mark=*,dashed] table [x=contrast, y=bestapprox-L-infty-L2-error-u, col sep=comma]{../dataNew/simplexExact_MultContrast_T0.5.csv};

@@

                \addplot[black,line width=3pt,mark=None] table [x=x, y=y, col sep=comma] {../dataNew/plots/exact_plotSJ_contrast2.5_time0.25.csv};

@@

                \legend{exact,$L=1$, $L=2$, $L=3$, $L=4$}
                \nextgroupplot[width=9cm,height=7cm,domain=0:1,xmode=linear,ymode=linear, xlabel={$x$}, ylabel={}, title={$u$ vs. $u_h$ ($c_1 = 5.5$)}, %cycle list name=paulcolors, 
                legend pos=north west, %yticklabels={,,},
                ymin = -2.25, ymax = 2.25,
                ]
                %\draw[fill=gray!40,draw=none,opacity=0.5] (0.25,-3) -- (0.25,3) -- (0,3) -- (0,-3) -- cycle;
                %\draw[fill=gray!40,draw=none,opacity=0.5] (0.75,-3) -- (0.75,3) -- (1.0,3) -- (1.0,-3) -- cycle;
                %\addplot+[discard if not={contrast}{1.0},line width=2.5pt,mark=None,color=black] table [x=x, y=y, col sep=comma] {../../wave_data_assimilation/data/jumpCoefs/exact_plot_data.csv};

                \addplot[black,line width=3pt,mark=None] table [x=x, y=y, col sep=comma] {../dataNew/plots/exact_plotSJ_contrast5.5_time0.25.csv};

@@

            \end{groupplot}
        \end{tikzpicture}
        \pgfplotslegendfromname{named}
    \end{center}
    \caption{Exact solution \eqref{eq:1D:exact:simple} and approximations with $k = 3$ on refinement levels $L \in \{1,2,3,4 \}$ for contrasts $c_1 = 2.5$ (left) and $c_1= 5.5$ (right). The scale in the right plot indicates the mesh size $h = 1/2^{L+1}$ for the different refinement levels.v \textbf{generate the data for the exact plots, at which time are we? $t = 0.25$}}
    \label{fig:jumpCoefs:quality}
  \end{figure}

@@

                \addplot[black,line width=3pt,mark=None] table [x=x, y=y, col sep=comma] {../dataNew/plots/exact_plotMJ_contrast7.5_time0.25.csv};

                \addplot+[discard if not={L}{3},discard if not={contrast}{7.5},line width=1.0pt,mark=None,color=orange] table [x=x, y=y, col sep=comma] {../dataNew/plots/MultipleJumpsExact_ApproxPlot_T0.5_L3_contrast7.5.csv};

                \addplot+[discard if not={L}{4},discard if not={contrast}{7.5},line width=1.0pt,mark=None,color=blue!75!white] table [x=x, y=y, col sep=comma] {../dataNew/plots/MultipleJumpsExact_ApproxPlot_T0.5_L4_contrast7.5.csv};
